Investigating soundscapes perception through acoustic scenes simulation

### 2018-07-23 lafaySimulatedScenePerception.tex
nm comments

    \address{$^{1)}$ Laboratoire des Sciences du Numérique de Nantes-CNRS-\'Ecole Centrale de Nantes, Nantes, France.\\
    \hspace*{8pt}mathieu.lagrange@cnrs.fr\\
    $^{2)}$ STMS Ircam-CNRS-UPMC Institut de Recherche et Coordination Acoustique/Musique, Paris, France}

@@

In order to account for the possibility that the morphological differences between soud events and sound textures may have some important consequences on the perception of the scenes, the simulation tool \emph{Simscene} follows this distinction and considers two distinct sound databases: one with classes of events only and the other with classes of textures only. These two types of sound classes also have specific scheduling procedures during the simulation process.

@@

\begin{figure}[t]
        \myfloatalign
        \includegraphics[width=.8\linewidth]{gfx/3.pdf}
       \caption{Hierarchical organisation of the isolated sounds used in the simulation.}\label{fig:orgDb}
\end{figure}

@@

Once the typology and the set of sounds are available, an important design issue is the need of a suitable way to display the sound dataset to the user. Most browsing tools are based on keyword indexing; however, for sensory experiments that study the objectivation of a subject's mental representations, this means is problematic: indeed, the availability of a verbal description of the sound can influence the subject's choice, and potentially induce biases in the analysis conclusions. For example, a subject may automatically select sounds referenced as belonging to a \emph{park} environment to build a \emph{calm} soundscape, rather than focusing on their perception.

@@

44 students (30 male, 14 female; averaging $21.6$ years of age, s.d. of $2.0$ years) from \emph{Ecole Centrale de Nantes} (a french engineering school) took part in the experiment. All the subjects had been living in Nantes, France, for at least two years at the time of the experiment.

@@

Due to temporal constraints, subjects only assess 30 seconds of the initial 1-minute simulated scenes generated in experiment 1.a (from second 15 to sec. 45).

@@

10 students (8 male, 2 female; averaging $23.1$ years of age, s.d. of $1.8$ years) from \emph{Ecole Centrale de Nantes} took part to the experiment. All the subjects had been living in Nantes, France, for at least two years at the time of the experiment. None of them took part to the previous simulation experiment (experiment 1.a).

@@

If the null hypothesis is rejected, the class $j$ is said to be typical with respect to the type of scene $k$. Such typical classes are called \textbf{sound markers}. Testing is done for each class, at each level of abstraction, and separately for texture and event classes.

@@

\begin{figure}[t]
		\myfloatalign
        \subfloat[]
        {\includegraphics[width=.33\linewidth]{gfx/xp_soundlevel_1}\label{fig:soundlevela}}
        \subfloat[]
        {\includegraphics[width=.33\linewidth]{gfx/xp_soundlevel_3}\label{fig:soundlevelb}}
        \subfloat[]
        {\includegraphics[width=.33\linewidth]{gfx/xp_soundlevel_5}\label{fig:soundlevelc}}\par
        \subfloat[]
        {\includegraphics[width=.33\linewidth]{gfx/xp_soundlevel_2}\label{fig:soundleveld}}
        \subfloat[]
        {\includegraphics[width=.33\linewidth]{gfx/xp_soundlevel_4}\label{fig:soundlevele}}
        \subfloat[]
        {\includegraphics[width=.33\linewidth]{gfx/xp_soundlevel_6}\label{fig:soundlevelf}}
       \caption{Distributions of the sound levels $L$ (a, d), $L(E)$ (b, e) et $L(T)$ (c, f), with respect to scene type (i: ideal, ni: non-ideal) (a, b, c) and perceived pleasantness $\mathcal{A}_{scene}$ of experiments 1.b (d, e, f).}
\end{figure}

@@

\begin{enumerate}
\item a significant difference between levels of ideal and non-ideal scenes exists ($L_m$, $L(E)_m$ et $L(T)_m$: $p<0.01$);
\item the sound level of scenes is mainly related to the sound events, compared to the textures;
\item the sound level of events has an influence on the perception of pleasantness for non-ideal scenes, but not for ideal scenes;
\item the sound level of textures does not play any role in the perception of the pleasantness.
\end{enumerate}

@@

12 subjects perform the test (8 male, 4 female; averaging $29.5$ years of age, s.d. of $14$ years). None of them have performed experiments 1.a and 1.b. All the subjects had been living in Nantes, France, for at least two years at the time of the experiment. All the subjects performed the test successfully.

@@

\begin{itemize}
\item The simulator can be deployed on a large scale via the Internet thanks to the web-based software architecture;
\item Simulated scenes can be analyzed without the need to take into account the different mother tongues of the subjects, the semantic nature of the classes of sounds used being known \emph{a priori} by the experimenter and without the need to analyze and annotate the sound scene to identify the sources, their occurrences in the simulated scene being directly available.
\end{itemize}

>>>>

    \address{$^{1)}$ Laboratoire des Sciences du Numérique de Nantes-CNRS-\'Ecole Centrale de Nantes, Nantes, France.\\
    \hspace*{8pt}mathieu.lagrange@cnrs.fr\\
    $^{2)}$ STMS Ircam-CNRS-SU Institut de Recherche et Coordination Acoustique/Musique, Paris, France}

@@

In order to account for the possibility that the morphological differences between sound events and sound textures may have some important consequences on the perception of the scenes, the simulation tool \emph{Simscene} follows this distinction and considers two distinct sound databases: one with classes of events only and the other with classes of textures only. These two types of sound classes also have specific scheduling procedures during the simulation process.

@@

\begin{figure}[t]
        \myfloatalign
        \includegraphics[width=\linewidth]{gfx/3.pdf}
       \caption{Hierarchical organisation of the isolated sounds used in the simulation.}\label{fig:orgDb}
\end{figure}

@@

Once the typology and the set of sounds are available, an important design issue is the need of a suitable way to display the sound dataset to the user. Most browsing tools are based on keyword indexing; however, for sensory experiments that study the objectivation of a subject's mental representations, this can be problematic as the availability of a verbal description of the sound can influence the subject's choice, and potentially induce biases in the analysis conclusions. For example, a subject may automatically select sounds referenced as belonging to a \emph{park} environment to build a \emph{calm} soundscape, rather than focusing on their perception.

@@

44 students (30 male, 14 female; averaging $21.6$ years of age, s.d. of $2.0$ years) from \emph{Ecole Centrale de Nantes} (a french engineering school) took part in the experiment. All the subjects had been living in Nantes, France, for at least two years at the time of the experiment and reported normal hearing.

@@

Due to temporal constraints, subjects only assess 30 seconds of the initial 1-minute simulated scenes generated in experiment 1.a (from sec. 15 to sec. 45).

@@

10 students (8 male, 2 female; averaging $23.1$ years of age, s.d. of $1.8$ years) from \emph{Ecole Centrale de Nantes} took part to the experiment. All the subjects had been living in Nantes, France, for at least two years at the time of the experiment and reported normal hearing. None of them took part to the previous simulation experiment (experiment 1.a).

@@

If the null hypothesis is rejected, the class $j$ is said to be typical with respect to the type of scene $k$. Such typical classes are called \textbf{sound markers}, in reference to the work of \cite{schafer1993soundscape}. Testing is done for each class, at each level of abstraction, and separately for texture and event classes.

@@

\begin{figure}[t]
		\myfloatalign
        \subfloat[]
        {\includegraphics[width=.33\linewidth]{gfx/xp_soundlevel_1}\label{fig:soundlevela}}
        \subfloat[]
        {\includegraphics[width=.33\linewidth]{gfx/xp_soundlevel_3}\label{fig:soundlevelb}}
        \subfloat[]
        {\includegraphics[width=.33\linewidth]{gfx/xp_soundlevel_5}\label{fig:soundlevelc}}\par
        \subfloat[]
        {\includegraphics[width=.33\linewidth]{gfx/xp_soundlevel_2}\label{fig:soundleveld}}
        \subfloat[]
        {\includegraphics[width=.33\linewidth]{gfx/xp_soundlevel_4}\label{fig:soundlevele}}
        \subfloat[]
        {\includegraphics[width=.33\linewidth]{gfx/xp_soundlevel_6}\label{fig:soundlevelf}}
       \caption{Distributions of the sound levels $L$ (a, d), $L(E)$ (b, e) and $L(T)$ (c, f), with respect to scene type (i: ideal, ni: non-ideal) (a, b, c) and perceived pleasantness $\mathcal{A}_{scene}$ of experiments 1.b (d, e, f).}
\end{figure}

@@

\begin{enumerate}
\item a significant difference between levels of ideal and non-ideal scenes exists ($L_m$, $L(E)_m$ and $L(T)_m$: $p<0.01$);
\item the sound level of scenes is mainly related to the sound events, compared to the textures;
\item the sound level of events has an influence on the perception of pleasantness for non-ideal scenes, but not for ideal scenes;
\item the sound level of textures does not play any role in the perception of the pleasantness.
\end{enumerate}

@@

12 subjects perform the test (8 male, 4 female; averaging $29.5$ years of age, s.d. of $14$ years). None of them have performed experiments 1.a and 1.b. All the subjects had been living in Nantes, France, for at least two years at the time of the experiment and reported normal hearing. All the subjects performed the test successfully.

@@

\begin{itemize}
\item The simulator can be deployed on a large scale via the Internet thanks to the web-based software architecture, provided that the sound databases are adapted to the cultural background of the cohort;
\item Simulated scenes can be analyzed without the need to take into account the different mother tongues of the subjects, the semantic nature of the classes of sounds used being known \emph{a priori} by the experimenter and without the need to analyze and annotate the sound scene to identify the sources, their occurrences in the simulated scene being directly available.
\end{itemize}
